### 2020-07-04 abstract.tex
final edits for 2nd draft

This thesis presents the Molten Salt Fast Reactor simulation capabilities of
Moltres, a simulation tool for molten salt reactors.

>>>>

Molten salt reactors are a class of advanced nuclear reactors that promise
numerous improvements over the current fleet of largely light-water reactors.
As the world continues its transition towards low-carbon electricity
generation to combat climate change, molten salt reactors are potential
options in the near-term future for replacing fossil fuel and ageing nuclear
power plants. At the current state of development, molten salt reactors still
require extensive research to become viable. This thesis presents latest
developments in Moltres, a simulation tool for molten salt reactors. These new
developments are: support for coupling the incompressible Navier-Stokes and
the delayed neutron precursor looping systems, and a new decay heat model for
simulating decay heat from fission products at steady state and during
transients. This work demonstrates these capabilities in conjunction with
existing capabilities through multiphysics simulations of the Molten Salt Fast
Reactor concept. This work first verifies the six-group neutron diffusion
results from Moltres against continuous-energy Monte Carlo neutron transport
results from Serpent 2. The multiplication factors $k_{\text{eff}}$, delayed
neutron fractions $\beta$, temperature reactivity coefficient $\alpha_T$, and
the six-group neutron energy spectra from Moltres agreed with the high
fidelity simulation results from Serpent 2. The $k_{\text{eff}}$ values have
small discrepancies on the order of 100 pcm, which is smaller than some
reported literature values. The decay heat model shows an expected flattening
of the temperature distribution due to the movement and dispersion of the
decay heat precursors throughout the primary coolant loop. This work also
demonstrates and verifies steady state and
transient multiphysics simulations of the Molten Salt Fast Reactor. The
transient scenarios under study are unprotected instances of reactivity
insertion, loss of heat sink, loss of flow, and pump overspeed. The steady
state and transient results are verified against results from another paper
that presented results for the same cases. The steady-state temperature and
velocity distributions, and the peak neutron flux showed good agreement with
the literature results. Minor differences in the delayed neutron precursor
distribution and the in-core delayed neutron fraction were explainable with
the differences in the handling of turbulence in the models. In three of the
transient results (reactivity insertion, loss of heat sink, and pump
overspeed), Moltres reproduced the expected magnitude and pattern of the
reactor response to these transient initiators.
The loss of flow results showed greater discrepancies that are attributed
to differences in the fluid dynamics modeling in Moltres and the other models.
Through the verification studies, this work has also identified avenues
for further Moltres software development.

### 2020-07-04 conclusion.tex
final edits for 2nd draft

The unique phenomena in \glspl{MSR}, generally not found in conventional
reactors, necessitate the development of new reactor safety analysis software.
This thesis presents the latest developments in Moltres, namely coupling its
existing neutron diffusion module to the incompressible Navier-Stokes module
in MOOSE, and developing a decay heat model for short-term transients. We
demonstrated and verified some of its current capabilities through a static
neutronics study, and a coupled neutronics/thermal-hydraulics safety analysis
of the \gls{MSFR} concept.

@@

The unprotected reactivity insertion and loss of heat sink results showed the
same trends Fiorina et al. observed in their Polimi and TUDelft models. The
small difference in the temperature reactivity coefficient accounted for the
small difference in the magnitude of the peaks in power output and average
core temperature increase. The differences between the pointwise heat
exchanger in Moltres and the volumetric heat exchanger in the other two
models required minor adjustments in the relationship between flow rate
$\dot{m}$ and the heat transfer coefficient $h$ from the original
Dittus-Boelter correlation for the pump-initiated transients. Assuming a
directly proportional relationship between $\mu_t$ and $\dot{m}$ yielded
results in good agreement with the other two models for the pump overspeed
transient. However, Moltres performed poorly in the loss of flow transient as
we could not incorporate buoyancy-driven flow and its associated effects on
$\mu_t$.

@@

Further research and development on Moltres should aim to rectify the issues
mentioned in this thesis. There are three main avenues for improvement.
Firstly, proper 2D/3D heat exchanger implementation
would allow us to move away from the 1D outer loop system and towards a full
2D/3D closed loop. The biggest change from this is users would be able to use
the Boussinesq approximation for buoyancy-driven flow capability in Moltres.
Buoyancy-driven flow is a critical component in loss of forced flow scenarios
and these scenarios in turn are important accident transients in reactor
safety analyses.

>>>>

The unique phenomena in \glspl{MSR}
necessitate the development of new reactor safety analysis software.
This thesis presents the latest developments in Moltres, namely coupling its
existing neutron diffusion module to the incompressible Navier-Stokes module
in MOOSE, and developing a decay heat model for short-term transients. We
demonstrated and verified some of its current capabilities through a static
neutronics study, and a coupled neutronics/thermal-hydraulics safety analysis
of the \gls{MSFR} concept.

@@

The unprotected reactivity insertion and loss of heat sink results showed the
same trends Fiorina et al. observed in their Polimi and TUDelft models. The
small difference in the temperature reactivity coefficient accounted for the
small difference in the magnitude of the peaks in power output and average
core temperature increase. The differences between the pointwise heat
exchanger in Moltres and the volumetric heat exchanger in the other two
models required minor adjustments in the relationship between flow rate
$\dot{m}$ and the heat transfer coefficient $h$ from the original
Dittus-Boelter correlation for the pump-initiated transients. Assuming a
directly proportional relationship between $\mu_t$ and $\dot{m}$ yielded
results in good agreement with the other two models for the pump overspeed
transient. However, Moltres performed poorly in the loss of flow transient as
this work could not incorporate buoyancy-driven flow and its associated
effects on $\mu_t$.

@@

Further research and development on Moltres should aim to rectify the issues
mentioned in this thesis. This work has highlighted three main avenues for
improvement. Firstly, proper 2D/3D heat exchanger implementation
would allow us to move away from the 1D outer loop system and towards a full
2D/3D closed loop. The biggest change is that users would be able to use
the Boussinesq approximation for buoyancy-driven flow capability in Moltres.
Buoyancy-driven flow is a critical component in loss of forced flow scenarios
and these scenarios in turn are important accident transients in reactor
safety analyses.

### 2020-07-04 introduction.tex
final edits for 2nd draft

Greenhouse gas emission from human activities is the main cause of climate
change, a global phenomenon characterized by shifting weather patterns and
rising temperatures. The concentration of CO$_2$, a greenhouse gas, in the
atmosphere has been rising steadily throughout modern human history
\cite{us_department_of_commerce_global_nodate}. Climate change has dire
consequences on human health and safety due to extreme weather events and the
overall impact on food production \cite{noauthor_global_nodate}. As a result,
urgent measures are necessary to limit greenhouse gas emissions, namely CO$_2$
emissions from electricity generation.

Electricity generation from burning fossil fuels represents the
largest source of CO$_2$ emissions (38\% in 2018 \cite{iea_global_2019});
replacing it with low-carbon
alternatives would curb a large fraction of emissions. Nuclear power is
a viable low-carbon replacement for burning fossil fuels.
Other low-carbon energy sources include solar-, wind-, and hydro-power. Each
energy source has its own set of advantages and disadvantages; nuclear power's
advantage over renewable energy sources is that it provides consistent
base-load power independent of weather and geographical location
\cite{petti_future_2018}. Employing a diverse mix of these energy sources
ensures energy security and reliability in our transition
towards a low-carbon future \cite{petti_future_2018}.

The world would have to ramp up the current rate of reactor deployments to
displace some of the
presently large share of energy production from fossil fuel power plants.
However, several obstacles stand in the way of mass reactor deployments. These
include perceived safety risks, sustainability concerns, nuclear proliferation
risks, and the ability to compete economically with other sources of energy
\cite{massachusetts_institute_of_technology_future_2003}. A potential solution
to the aforementioned issues is the \gls{MSR} concept, one of six advanced
reactor designs selected by the Generation IV International Forum
\cite{gif_technology_2002} for continued research and development.

@@

Most reactor analysis applications are usually reactor-specific by
design such as PARCS and TRACE for \glspl{LWR}, and SAS4A/SASSYS-1 for
liquid metal cooled reactors. Thus, these applications generally would not
account for \gls{MSR}-specific phenomena and are not suitable for \gls{MSR}
analysis without any modifications to the source code. Some research efforts
do focus on adapting these applications for \gls{MSR} analysis. Examples
include the coupling of modified versions of TRACE and PARCS
\cite{pettersen_coupled_2016}, and the development of VERA-MSR from the
integrated \gls{LWR} simulation tool VERA \cite{graham_development_2019}.
Others developed their \gls{MSR} simulation tools from general
multiphysics or \gls{CFD} applications such as COMSOL
\cite{fiorina_modelling_2014} and OpenFOAM \cite{aufiero_development_2014}.

@@

This thesis demonstrates latest capabilities of Moltres
\cite{lindsay_introduction_2018}, an open source MSR simulation tool.
In particular, this thesis presents two more recent
developments in Moltres, namely fully integrating \gls{MOOSE}'s incompressible
Navier-Stokes module into Moltres, and introducing a
decay heat model.
The main objective of this thesis is to verify Moltres'
latest capabilities in modeling multiphysics, steady-state, and transient
behavior of fast-spectrum \glspl{MSR} through the study of the \gls{MSFR}
concept. Code-to-code verification is an important exercise in software
development for ensuring that the application produces accurate and reliable
results. This thesis covers the \gls{MSFR} concept mainly because 
there is readily available literature data to verify our results against. The
steady-state and transient results also feature a significant degree of
physics interaction between the neutronics and thermal-hydraulics. This
present work will first present a verification of Moltres' \gls{MSFR}
neutronics results against Serpent 2, followed by a verification of
the coupled neutronics/thermal-hydraulics steady-state and transient accident
results against results from the Polimi and TUDelft models published by
Fiorina et al. \cite{fiorina_modelling_2014}. The
secondary objective is to identify areas of improvement in Moltres for future
research.

>>>>

Greenhouse gas emission from human activities is the main cause of climate
change, a global phenomenon characterized by shifting weather patterns and
rising temperatures. The concentration of CO$_2$, a greenhouse gas, in the
atmosphere has been rising steadily throughout modern human history
\cite{us_department_of_commerce_global_2020}. Climate change has dire
consequences on human health and safety due to extreme weather events and the
overall impact on food production \cite{mcmichael_global_2004}. As a result,
urgent measures are necessary to limit greenhouse gas emissions, namely CO$_2$
emissions from electricity generation.

Electricity generation from burning fossil fuels represents the
greatest source of CO$_2$ emissions (38\% in 2018 \cite{iea_global_2019});
replacing it with low-carbon
alternatives would curb a significant fraction of emissions. Nuclear power is
a viable low-carbon replacement for burning fossil fuels.
Other low-carbon energy sources include solar-, wind-, and hydro-power. Each
energy source has its own set of advantages and disadvantages; nuclear power's
advantage over renewable energy sources is that it provides consistent
base-load power independent of weather and geographical location
\cite{petti_future_2018}. Employing a diverse mix of these energy sources
ensures energy security and reliability in our transition
towards a low-carbon future \cite{petti_future_2018}.

The world would have to ramp up the current rate of reactor deployments to
displace some of the presently large share of energy production from fossil
fuel power plants. However, several obstacles stand in the way of mass reactor
deployments. These obstacles include perceived safety risks, sustainability
concerns, nuclear proliferation
risks, and the ability to compete economically with other sources of energy
\cite{massachusetts_institute_of_technology_future_2003}. A potential solution
to the aforementioned issues is the \gls{MSR} concept, one of six advanced
reactor designs selected by the Generation IV International Forum
\cite{gif_technology_2002} for continued research and development.

@@

Most reactor analysis applications are usually reactor-specific by
design such as PARCS and TRACE for \glspl{LWR}, and SAS4A/SASSYS-1 for
liquid metal cooled reactors. Thus, these applications would disregard
\gls{MSR}-specific phenomena and are inappropriate for \gls{MSR}
analysis without any modifications to the source code. Some research efforts
do focus on adapting these applications for \gls{MSR} analysis. Examples
include the coupling of modified versions of TRACE and PARCS
\cite{pettersen_coupled_2016}, and the development of VERA-MSR from the
integrated \gls{LWR} simulation tool VERA \cite{graham_development_2019}.
Others developed their \gls{MSR} simulation tools from general
multiphysics or \gls{CFD} applications such as COMSOL
\cite{fiorina_modelling_2014} and OpenFOAM \cite{aufiero_development_2014}.

@@

This thesis demonstrates latest capabilities of Moltres
\cite{lindsay_introduction_2018}, an open source MSR simulation tool.
In particular, this thesis presents two more recent
developments in Moltres, namely fully integrating \gls{MOOSE}'s incompressible
Navier-Stokes module into Moltres, and introducing a
decay heat model.
The main objective of this thesis is to verify Moltres'
latest capabilities in modeling multiphysics, steady-state, and transient
behavior of fast-spectrum \glspl{MSR} through the study of the \gls{MSFR}
concept. Code-to-code verification is an important exercise in software
development for ensuring that the application produces accurate and reliable
results. This thesis covers the \gls{MSFR} concept mainly because it has been
studied extensively with readily available literature data to verify our
results against. The
steady-state and transient results also feature a significant degree of
physics interaction between the neutronics and thermal-hydraulics. This
present work will first present a verification of Moltres' \gls{MSFR}
neutronics results against Serpent 2, followed by a verification of
the coupled neutronics/thermal-hydraulics steady-state and transient accident
results against results from the Polimi and TUDelft models published by
Fiorina et al. \cite{fiorina_modelling_2014}. The
secondary objective is to identify areas of improvement in Moltres for future
research.
